Practical Type Blurring

### 2020-01-08 paper/paper.tex
signature blurring

We show that the techniques used for type reconstruction can be bypassed,
and we claim that the bypass is always possible given \emph{semantics gap} between
type systems of high-level and machine language.
% \keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}
Binary code type inference is to restore high-level type information of variables,
functions from machine codes. It is useful for reverse code engineering (e.g.
decompilation) or static binary rewriting/instrumentation, just to name a few.
Type reconstruction is available in several binary analysis
tools~\cite{noauthor_hex-rays_nodate, noauthor_jeb_nodate, noauthor_ghidra_nodate}
as a result of machine code decompilation. Academic research includes~\cite{mycroft_type-based_1999}
which is one of the pilot papers on this domain, an improvement supporting
a limited form of subtyping is presented in~\cite{lee_tie_2011}, a survey of
research up until 2016 is given in~\cite{caballero_type_2016}. One of recent
directions is to use machine learning~\cite{maier_typeminer_2019}, which is out of scope of
the paper. All approaches tend to output a C-like type system.

@@

Though most papers agree that type information is removed in compilation~\cite{lee_tie_2011,caballero_type_2016,lin_automatic_2010},
there is no explicit explication of \emph{why and how} some type information is not needed
(so removed) in machine codes. A popular point of view that types should attach with
storage specific primitives (e.g. \texttt{int} is some \texttt{32-bit} signed integer
stored in a register or stack), though mostly correct, does not reveal the nature of
types. This is one of the reasons why some researchers unify data structure reverse
engineering with type reconstruction~\cite{caballero_type_2016, caballero_polyglot_2007}.
In some languages other than C, types can be zero-sized~\cite{noauthor_phantom_nodate}, another example is
\emph{regions}~\cite{grossman_region-based_2002} which have no storage imprint.

@@

\paragraph{Contribution}
We first discuss the type information loss in compilation and some cases
where assigning a high-level \emph{type} for a low-level primitive
is not possible. Though this part is not novel, we bring it to the context
of binary code analysis to show limits of type reconstruction. Next, we present
a C compiler which tries to hide types from reconstruction techniques described
in current researches.

>>>>

We show that the techniques used for type reconstruction can be bypassed,
and we claim that the bypass is always possible given \emph{semantics gap} between
type systems of high-level and machine language.
% \keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}
Binary code type inference is to restore high-level type information of variables,
functions from machine codes. It is useful for reverse code engineering (e.g.
decompilation) or static binary rewriting/instrumentation, just to name a few.
Type reconstruction is available in several binary analysis
tools~\cite{noauthor_hex-rays_nodate, noauthor_jeb_nodate, noauthor_ghidra_nodate}
as a result of machine code decompilation. Academic research includes~\cite{mycroft_type-based_1999}
which was probably the pilot paper on the domain, an extension supporting
a limited form of subtyping is presented in~\cite{lee_tie_2011}, a survey of
research up until 2015 is given in~\cite{caballero_type_2016}. Recent
directions are using machine learning~\cite{maier_typeminer_2019} or statistical
language model~\cite{katz_estimating_2016}. All approaches tend to output a C-like type system.

@@

This bird view about binary code type reconstruction may give an impression that the
research domain has been well established already, the incorrectness comes from foreign
reasons: binary code disassembly, high-level variable detection, etc.
But there are more intrinsic problems, as we claim below.

\emph{First}, there is a common agreement that type information is removed through compilation~\cite{lee_tie_2011,caballero_type_2016,lin_automatic_2010},
but there is no explicit explication of why and how some type information is not needed
(so removed) in machine codes, actually not all type information can be restored.

\emph{Second}, while some researchers unify data structure reverse engineering with
type reconstruction~\cite{caballero_type_2016, caballero_polyglot_2007}, types do not
always attach with storage specific primitives (e.g. \texttt{int} is some \texttt{32-bit}
signed integer stored in a register or stack). The storage-based point of view though
mostly correct in low-level languages like C, does not reveal the nature of types. In
some languages other than C, types can be zero-sized~\cite{noauthor_phantom_nodate},
another example is \emph{regions}~\cite{grossman_region-based_2002} which have no
storage imprint.

@@

\paragraph{Contribution}
We first discuss the type information loss in compilation and some cases
where assigning a high-level \emph{type} for a low-level primitive
is not possible. Though this part is not novel, we bring it to the context
of binary code analysis to show limits of type reconstruction. Next, we present
a proof-of-concept C compiler which tries to hide types from reconstruction techniques described
in current researches.
